---
title: "Median income in PA counties"
author: "Steve R."
output:
  html_document:
    toc: yes
    toc_float: yes
description: PA median income choropleth
---

# Mapping median income by county in Pennsylvania ###

This is a simple analysis producing a data table and static choropleth graphic depicting the median income of Pennslyvania's counties, according to 2016 U.S. Census data.

### Install packages and load libraries
```{r Packages and libraries}
options(repos="https://cran.rstudio.com")
install.packages("rmarkdown")
install.packages("tidyverse")
install.packages("sf")
install.packages("viridis")
install.packages("tigris")
install.packages("ggplot2")
library(ggplot2)
library(sf)
library(tigris)
library(dplyr)
```

### Set sf option
```{r sf option}
options(tigris_class = "sf")
```

### Load Pennsylvania data
```{r load state data}
pa <- counties("PA", cb=T)
```

### Plot Pennsylvania counties
```{r Plot Pennsylvania geograpy}
ggplot(pa) + 
  geom_sf() +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) +
  labs(title="Pennsylvania counties")
```

### Add Census key
```{r Census key}
Sys.setenv(CENSUS_KEY="bb35319b92c113f304adfb661e1b0ada4e7b5526")
readRenviron("~/.Renviron")
Sys.getenv("CENSUS_KEY")
```

### Install censusapi package and load library
```{r censusapi}
install.packages("censusapi")
library(censusapi)
```

### Load Pennsylvania income data
```{r Load income data}
pa_income <- getCensus(name = "acs/acs5", vintage = 2016, 
                       vars = c("NAME", "B19013_001E", "B19013_001M"), 
                       region = "county:*", regionin = "state:42")
head(pa_income)
```

### Join pennsylvania geography and income data
```{r Join income and geography data}
pa4ever <- left_join(pa, pa_income, by=c("COUNTYFP"="county"))
View(pa4ever)
```

### Plot 2016 median income by county in Pennsylvania
```{r Plot county income data}
ggplot(pa4ever) + 
  geom_sf(aes(fill=B19013_001E), color="white") +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) +
  scale_fill_distiller(palette="Greens", direction=1, name="Median income", labels=scales::dollar_format()) +
  labs(title="2016 Median income in Pennsylvania counties", caption="Source: US Census/ACS5 2016")
```

### View table showing median income by county in Pennsylvania
```{r View data table}
County <- pa4ever$NAME.x
Median_Income <- paste("$",format(pa4ever$B19013_001E, big.mark=","),sep="")
med_income <- cbind(County,Median_Income)
library(DT)
datatable(med_income,extensions = "Buttons", options = list(
  dom = "Bfrtip",
  buttons = c("copy", "csv", "excel", "pdf", "print")))
```
